Network Configuration Synthesis with Abstract Topologies

### 2016-11-14 paper/abstraction/paper.tex
edits for section 6

\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/templates}
  \caption{Example of a small configuration component of a router template for the datacenter spine role.}
  \label{fig:templates}
  \vspace{-1em}
\end{figure}

@@

%While templates help avoid a class of configuration errors (e.g., inconsistent routing adjacencies), they provide no guarantees
As described earlier (\S\ref{sec:introduction}), templates are hard to author and hard to validate. At the same time, templates that work for one topology may not work for seemingly-inconsequential variations (which may arise after the network evolves).
Consider the network in Figure~\ref{fig:example2}, which is similar to Figure~\ref{fig:example}: it has the same five roles, connected in a similar hierarchy. One might think that the same templates (with different database entries) can be used for both cases. However, if the templates are configured to disallow ``valley" paths (per policy (5) above), they will work for Figure~\ref{fig:example} but silently violate the fault tolerance policy ($(8)$) when used for Figure~\ref{fig:example2}.  Specifically, in Figure~\ref{fig:example2}, an aggregation-based blackhole (\S\ref{sec:background}) will occur when the link S1--A1 fails; after this failure, S1 has no valley-free path to P[1--2] even though it will continue to get traffic for these prefixes as it announces the covering prefix PG (because it gets routes for P[3--4]). No such blackhole will occur in Figure~\ref{fig:example} after a single link failure because spine routers have two links to each pod.

@@

  \vspace*{1em}
  \begin{minipage}[t]{\linewidth}%
  \begin{tikzpicture}
    % \inference{0}{L$(j,k)$}{$e_i > 0$}{S$(min(j*k,g),1)$}{I-out3}{$e_1$}{$e_2$}
    \inferencetwo{0}
    \node[anchor=west] at (1,2.2) {where $g=n-\dfrac{(m-j)*e_1}{e_2}$};
    \inferencelocal{6.5}{.4}
    \inferenceglobal{12}{.4}
  \end{tikzpicture}
  \end{minipage}

@@

The first rule, \infrule{I-out1} applies to a learned invariant of the form $L_m(j,k)$ where the number of outgoing edges from any concrete node in the $m$ role is greater than 0. In the worst case, the largest group of concrete nodes we could hope to reach at the $n$ role would be $e_1$ since all j nodes at the bottom may have outgoing edges to the same concrete nodes at the top. Furthermore, the total number of disjoint paths to the $j$ nodes at the bottom is equal to $j*k$. Since adding disjoint edges between the nodes keeps the existing paths disjoint, and since we can not exceed the current number of disjoint paths to the concrete nodes in role $m$ on the bottom, the largest reachable group for the role $n$ on the top will be $min(j*k,e_1)$. However, because we do not know for certain that all $j$ nodes in role $m$ have outgoing edges to the same nodes at the top, we conservatively use $1$ for the number of disjoint paths to each node on top.

Consider rule \infrule{I-out2}, and consider any node in the role $n$. There are $e_2$ incoming edges to that node. Due to the fact that A$_m(j,k)$, we know that at least j of those $e_2$ edges are connected to nodes with disjoint paths from the origin. Hence we infer A$_n(1,min(j,e_2))$.

@@

The rule \infrule{I-striping} is the most complicated case.
It starts with the invariant L$_m(j,k)$ at role $m$ and can be applied if each edge multiplicity $e_i > 0$ is valid given the constraints. The first inference for role $n$ tries to find the largest reachable group with disjoint paths to each. The idea is similar to the rule \infrule{I-out1}, but is able to use the fact that $e_2 > 0$ to learn more about the structure of the concrete topology. In particular it uses the following inequality, where $g$ represents the size of the group for the role $n$:
%
\[ \begin{array}{c}
  (m-j)*e_1 \geq (n-g)*e_2
\end{array} \]
\noindent
%
The remaining nodes ($m-j$) that are not part of the reachable group in the bottom role, each have $e_1$ outgoing edges and must be able to at least ``fill" the incoming edges for the remaining nodes not in the reachable group at the top ($n-g$) with $e_2$ incoming edges each. Solving the inequality gives the lower bound for $g$ used in Figure~\ref{fig:inference-rules}.

@@

Finally, rules \infrule{I-local} and \infrule{I-global} allow us to reason across pod hierarchies. The first rule, \infrule{I-local}, says that if there is an inference from L$_m(j,k)$ to L$_n(j',k')$ on the left, then if both nodes are in the same local pod P, we can leave all labels up to P unchanged. The global rule \infrule{I-global} says that when the edge goes between pods, we can infer the fact for all pods Q since the multiplicities apply uniformly for each pod.

>>>>

\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/templates}
  \caption{Idealized small configuration component for the datacenter spine role based on router templates from a large cloud provider (\IE, macros over gobbledygook).}
  \label{fig:templates}
  \vspace{-.8em}
\end{figure}

@@

%While templates help avoid a class of configuration errors (e.g., inconsistent routing adjacencies), they provide no guarantees
As described earlier (\S\ref{sec:introduction}), templates are hard to author and hard to validate. Worse, templates that work for one topology may not work for seemingly-inconsequential variations (which may arise after the network evolves).
Consider the network in Figure~\ref{fig:example2}, which is similar to Figure~\ref{fig:example}: it has the same five roles, connected in a similar hierarchy. One might think that the same templates (with different database entries) can be used for both cases. However, if the templates are configured to disallow ``valley" paths (per policy (5) above), they will work for Figure~\ref{fig:example} but silently violate the fault tolerance policy ($(8)$) when used for Figure~\ref{fig:example2}.  Specifically, in Figure~\ref{fig:example2}, an aggregation-based blackhole (\S\ref{sec:background}) will occur when the link S1--A1 fails; after this failure, S1 has no valley-free path to P[1--2] even though it will continue to get traffic for these prefixes as it announces the covering prefix PG (because it gets routes for P[3--4]). No such blackhole will occur in Figure~\ref{fig:example} after a single link failure because spine routers have two links to each pod.

@@

  \vspace*{1em}
  \begin{minipage}[t]{\linewidth}%
  \begin{tikzpicture}
    % \inference{0}{L$(j,k)$}{$e_i > 0$}{S$(min(j*k,g),1)$}{I-out3}{$e_1$}{$e_2$}
    \inferencetwo{0}
    \node[anchor=west] at (1,2.2) {where $g \geq n-\dfrac{(m-j)*e_1}{e_2}$};
    \inferencelocal{6.5}{.4}
    \inferenceglobal{12}{.4}
  \end{tikzpicture}
  \end{minipage}

@@

The first rule, \infrule{I-out1} applies to a learned invariant of the form $L_m(j,k)$ where the number of outgoing edges from any concrete node in the $m$ role is greater than 0. In the worst case, the largest group of concrete nodes we could hope to reach at the $n$ role would be $e_1$ since all j nodes at the bottom may have outgoing edges to the same concrete nodes at the top. Furthermore, the total number of disjoint paths to the $j$ nodes at the bottom is equal to $j*k$. Since extending the existing disjoint paths with disjoint edges keeps the paths disjoint, and since we can not exceed the current number of disjoint paths to the concrete nodes in role $m$ on the bottom, the largest reachable group for the role $n$ on the top will be $min(j*k,e_1)$. 
%
We conservatively use $1$ for the number of disjoint paths to each node in role $n$, since when $n$ is very large, all reachable nodes in role $m$ might only have a single edge to completely different nodes in $n$.
%However, because we do not know for certain that all $j$ nodes in role $m$ have outgoing edges to the same nodes at the top, we conservatively use $1$ for the number of disjoint paths to each node on top. 
%For example, if $n$ is very large, then each node in role $m$ may have only a single edge to each reachable node in role $n$.

Consider rule \infrule{I-out2}, and consider any node in the role $n$. There are $e_2$ incoming edges to that node. Due to the fact that $A_m(j,k)$, we know that at least j of those $e_2$ edges are connected to nodes with disjoint paths from the origin. Hence we infer $A_n(1,min(j,e_2))$.

@@

The rule \infrule{I-striping} is the most complicated case.
It starts with the invariant $L_m(j,k)$ at role $m$ and can be applied if each edge multiplicity $e_i > 0$ is valid given the constraints. The first inference for role $n$ tries to find the largest reachable group with disjoint paths to each. The idea is similar to the rule \infrule{I-out1}, but is able to use the fact that $e_2 > 0$ to learn more about the structure of the concrete topology. In particular it uses the following inequality, where $g$ represents the size of the group for the role $n$:
%
\[ \begin{array}{c}
  (m-j)*e_1 \geq (n-g)*e_2
\end{array} \]
\noindent
%
The remaining nodes ($m-j$) that are not part of the reachable group in the bottom role, each have $e_1$ outgoing edges and must be able to at least ``fill" the incoming edges for the remaining nodes not in the reachable group at the top ($n-g$) with $e_2$ incoming edges each. Solving the inequality gives the lower bound for $g$ used in Figure~\ref{fig:inference-rules}.

@@

Finally, rules \infrule{I-local} and \infrule{I-global} allow us to reason across pod hierarchies. The first rule, \infrule{I-local}, says that if there is an inference from $L_m(j,k)$ to $L_n(j',k')$ on the left, then if both nodes are in the same local pod P, we can leave all labels up to P unchanged. The global rule \infrule{I-global} says that when the edge goes between pods, we can infer the fact for all pods Q since the multiplicities apply uniformly for each pod.
